Overlapping Batches for the Assessment of Solution Quality in Stochastic Programs

### 2014-12-29 formatted/omrp_lb_acmtomacs.tex
Formatted According to ACM TOMACS
All the files, images, etc. are in the new subfolder 'formatted'

\documentclass[12pt]{article}

\usepackage{latexsym}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{enumerate}
\usepackage{multirow}
\usepackage{booktabs}
%\usepackage{mathptmx}
\usepackage[square]{natbib}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsbsy}
\usepackage{amsthm}
\usepackage[pdftex,colorlinks=true,urlcolor=blue,citecolor=black,anchorcolor=black,linkcolor=black]{hyperref}
\usepackage{setspace}

\onehalfspacing

\newtheorem{theorem}{Theorem}
\renewcommand{\thetheorem}{ \arabic{theorem}}
\newtheorem{corollary}[theorem]{Corollary}
\renewcommand{\thecorollary}{\arabic{corollary}}
\newtheorem{definition}{Definition}
\renewcommand{\thedefinition}{\arabic{definition}}
\newtheorem{lemma}{Lemma}
\renewcommand{\thelemma}{ \arabic{lemma}}
\newtheorem{proposition}{Proposition}
\renewcommand{\theproposition}{ \arabic{proposition}}
\newtheorem{remark}{Remark}
\renewcommand{\theremark}{ \arabic{remark}}

% Frequently used general mathematics
\newcommand{\R}{{\mathbb{R}}}
\newcommand{\Rp}{\R^+}
\newcommand{\Z}{{\mathbb{Z}}}
\newcommand{\Zp}{\Z^+}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\N}{\mathbb{N}}

@@

\newcommand{\Keywords}[1]{\par\noindent 
{\small{\em Keywords\/}: #1}}

%#########################################################
%*
%*  The Document.
%*
\begin{document}

\title{Overlapping Batches for the Assessment of Solution Quality in Stochastic Programs}

\author{David~Love$^{a}$\ \ and G\"{u}zin~Bayraksan$^{b}$\thanks{Corresponding author. Tel.: +1 614-292-1695. Fax:+1 614-292-7852}\\[6pt]
{\small
      $^{a}$ Program in Applied Mathematics, University of Arizona, Tucson, AZ 85721, USA,}\\
{\small \ \ \ \ \ \texttt{dlove@math.arizona.edu}} \\
{\small 
      $^{b}$ Integrated Systems Engineering, The Ohio State University, Columbus, OH 43210, USA,}\\
{\small \ \ \ \  \   \texttt{bayraksan.1@osu.edu}}}
\date{}

@@

\begin{abstract}
\noindent Overlapping Batch Means (OBM) has long been used in simulation as a method of reusing data to generate variance estimators with asymptotically lower variance.
In this paper, we apply the OBM method to stochastic programming by formulating a variant of the multiple replications procedure used for assessing solution quality.
We give conditions under which the resulting optimality gap point estimators are strongly consistent, the optimality gap interval estimators are asymptotically valid, and the OBM variance estimators for optimality gap have asymptotically lower variances relative to their nonoverlapping counterparts \citep{Meketon1984,Welch1987}.
We investigate computational efficiency, a combined measure of variance and computation time, providing guidelines on the degree of overlap.  
Numerical experiments on several test problems are presented, examining the small-sample behavior and the empirical computational efficiency of the overlapping batches method in this context.  
\medskip

\Keywords{Sample average approximation (SAA); Overlapping batch means;   Stochastic programming; Optimality gap estimation}
\end{abstract}

@@

The Overlapping Batch Means (OBM) method, proposed by \citet{Meketon1984}, is commonly used in simulation to obtain variance estimators having variances that are comparatively lower than their nonoverlapped counterparts. 
In the context of steady-state simulation, OBM is often used to estimate the variance of the sample mean, which is itself an estimator for the mean. 
\cite{SAH90} use overlapping estimators to estimate the variance of estimators for non-means. 
More recently, overlapping estimators based on standardized time series (as opposed to batch means) have been evaluated \citep{Alexopoulos01012007,Alexopoulos2007}; see also \cite{Meterelliyoz_etal_12}.
To the best of our knowledge, OBM has never before been studied in a stochastic optimization context. 
In this paper, we show that the overlapping batches achieve nearly the same bias and lower variances relative to their nonoverlapped counterparts when used in a stochastic optimization setting.

A natural setting where batch means are used in stochastic optimization (especially in stochastic programming) is for assessing solution quality. 
This typically involves generating point and interval estimators of the optimality gap of a given candidate solution using Monte Carlo simulation. 
Because optimality gap estimators can be difficult to analyze statistically, a common approach is to use nonoverlapping batches of data to generate a number of optimality gap estimators. 
These individual estimators are then used to form point and interval estimators applying the nonoverlapping batch means method.
This is referred to as the Multiple Replications Procedure (MRP) and was initially discussed in \cite{Mak1999}.   
In this paper, we formulate an overlapping batches version of MRP.

@@

We wish to assess the quality of a candidate solution $\xh \in X$, which may have been found in any way.
Assessing solution quality is important in practice because as (\ref{eq:sto_prog}) typically cannot be solved exactly, one only has an approximate solution $\xh$ without verification of its quality.
Assessing solution quality is also a critical component of stopping criteria in algorithms designed to (approximately) solve (\ref{eq:sto_prog}).
We define the quality of $\xh$ by its optimality gap, $\e{f(\xh,\xit)} - \zs$. 
The lower the optimality gap, the higher the quality of the solution; and a zero optimality gap implies that the solution is optimal. 
We assume that the candidate solution $\xh$ is given as input to our procedures, and hence fixed, throughout the paper.

The optimality gap $\e{f(\xh,\xit)} - \zs$ often cannot be calculated exactly. First, for a given candidate solution $\xh \in X$, evaluation of $\e{f(\xh,\xit)}$ typically involves a difficult multidimensional integral. 
Second, we do not know $\zs$. 
In many fields of optimization, the second problem is alleviated by obtaining lower bounds on $\zs$ through relaxations such as integrality, Lagrangian, or semidefinite relaxations.
As a result, it is common to evaluate an upper bound on the optimality gap. 
This motivates estimating the upper bound on the optimality gap in stochastic programming by using Monte Carlo simulation.
Given a sample size $m$, an upper bound on the optimality gap can be obtained by $$
\e{f(\xh,\xit)} - \zs \leq \e{f(\xh,\xit)} - \e{\zs_m}
$$ 
due to the inequality $\e{\zs_m} \leq \zs$ \citep{Mak1999,norkin_pflug_ruszczynski_98}.
This bound improves as the sample size increases; that is, $\e{\zs_m} \leq \e{\zs_{m+1}} \leq \zs$.  
A straightforward estimate of $\e{f(\xh,\xit)}$ is the sample mean, $\frac{1}{m} \sum_{i=1}^m f(\xh,\xiti)$. 
Instead of $\e{\zs_m}$, we can simply use $\zs_m$.  
The resulting point estimator of the optimality gap is then 
$$
\frac{1}{m} \sum_{i=1}^m f(\xh,\xiti) - \zs_m.
$$ 
Here, we assume the same observations $\xit^1, \dots, \xit^m$ are used in both terms.

Computing the above optimality gap estimate involves solving the optimization problem (\ref{eq:sto_prog_m}) to obtain a lower bound estimator of $\zs$, $\zs_m$, which complicates the statistical analysis. 
As mentioned above, to enable statistical inference, the MRP of \citet{Mak1999} generates $k$ independent estimators, each using sample size $m$ ($k$ nonoverlapping batches of size $m$), and averages them to obtain a point estimator.  
The sample variance of these estimators is used to form confidence intervals (CIs) for the optimality gap (we further review MRP in \S \ref{ssec:mrp}).  
An advantage of MRP is its applicability to a wide range of problems.  
With i.i.d.\ sampling, $f$ can be linear or nonlinear, $\X$ can include integrality constraints or not.  
It is also easy to implement; thus, MRP has been applied to a wide variety of problems, including finance, supply chain network design, etc.; see, e.g., \cite{bertocchi_etal_99,janjarassuk_linderoth_08,santoso_ahmed_etal_05}.  
Recently, the approach of (nonoverlapping) batching has been used for assessing solution quality of stochastic programs with finitely many expected value \citep{wang_ahmed_08} and stochastic dominance \citep{hu2010sample} constraints.

@@

The contributions of this article are as follows: 
\begin{itemize}
\item[(i)] We formulate an overlapping batch means variant of MRP, where the degree of overlap can be chosen by the user.  
This application of OBM is novel because each batch mean estimator includes an internal minimization, which has not been studied in the literature.
%\vspace*{-0.1in}

\item[(ii)] We provide conditions under which the resulting optimality gap estimators exhibit reduced variance when compared to their counterparts obtained via nonoverlapping batches.
We also show that the resulting point estimators are strongly consistent and that the interval estimators are asymptotically valid. 
These results require new proofs because of the presence of the optimization operator.

\item[(iii)] We examine small-sample behavior of the estimators through numerical examples. 
Our experiments indicate that the use of overlapping estimators results in gap estimators achieving lower variance than their nonoverlapping counterparts, along with comparable bias and interval coverage---for both small- and large-sample cases.

\item[(iv)] As the degree of overlap increases, the number of batches---and optimization problems solved---increases because we need to solve a sampling problem per batch. 
At the same time, variances are lowered, and when warm starting is available, solution time per batch can decrease, yielding efficiency gains. 
We explore this trade-off between computation time and lower variances.
Our results indicate that when warm starting is effective, overlapping estimators are computationally more efficient than those formed via nonoverlapping batches.
Based on this study, we provide guidelines on selecting the degree of overlap and demonstrate efficiency on a newsvendor problem. 
\end{itemize}

An earlier version of the paper appeared in \cite{love2011overlapping}. 
Most of the theoretical results in the current version of the paper are new, including almost sure convergence and asymptotically lower variances of the overlapping variance estimators, and asymptotic validity of the interval estimators. 
These results are achieved through a stronger assumption on the rate of convergence of the sampling problem's solution to the unique optimal solution, which is satisfied by several classes of (\ref{eq:sto_prog}); see \S \ref{subsec:assumptions}.
We also present additional computational results, including those from a large-scale problem, and study computational efficiency of the overlapping MRP method.

The rest of the paper is organized as follows.  
In the next section, we review relevant background information.  
We first discuss nonoverlapping and overlapping batch means in \S \ref{ssec:obm} and then briefly go over MRP in \S \ref{ssec:mrp}.  
In \S \ref{sec:omrp}, we present the Overlapping Multiple Replications Procedure (OMRP) with a parametrized degree of overlap.  
In \S \ref{sec:theory}, we show strong consistency of the OMRP point estimators, asymptotic validity of the OMRP CIs, and establish that the OMRP variance estimators have lower variances than their nonoverlapping MRP counterparts.
In \S \ref{sec:comp}, we test the performance of OMRP on four problems and study computational efficiency. 
We end in \S \ref{sec:concl} with a summary and conclusions.

@@

Consider a covariance stationary stochastic process $y^1, y^2, \dots$ with mean $\mu$ and variance parameter $0<\sigma^2<\infty$.  
The variance parameter is the sum of covariances at all lags, $\sigma^2=\mathrm{Var} \left(y^1 \right)+2\sum_{j=1}^{\infty}\mathrm{Cov} \left(y^1,y^{1+j}\right)$. 
If the variance parameter exists, then  $\sigma^2=\lim_{n\rightarrow \infty}n\mathrm{Var} \left(\ybb\right)$, where $\ybb=\frac{1}{n}\sum_{i=1}^{n} y^{i}$.
The task is to estimate $\mu$ given some realization of the univariate stochastic process $\y = (y^1, y^2, \dots, y^n)$.  
Typically, this involves forming a CI of the form $[L_\alpha(\y), U_\alpha(\y)]$ for a given level of significance $\alpha$ such that $\p{L_\alpha(\y) \leq \mu \leq U_\alpha(\y)} = 1 - \alpha$ or that this probability is approximately $1 - \alpha$ for large enough sample sizes.  
The usual estimator for $\mu$ is the sample mean $\ybb$, which is an unbiased estimator even in the presence of correlated data.  
However, the usual (sample) variance estimator $\frac{1}{n-1} \sum_{i=1}^n (y^i - \ybb)^2$ can be inappropriate to use in the presence of correlated data, resulting in inaccurate interval estimators.  
To overcome this difficulty, various variance estimators have been proposed in the simulation literature; see, e.g., \cite{law_15}.  
We briefly review two of these variance estimators: nonoverlapping and overlapping batch means.

@@

Nonoverlapping batch means takes $n$ observations $y^1, \dots, y^n$ of the stochastic process and splits them into $k$ batches of size $m$, where $k = \frac{n}{m}$ (for simplicity, assume, for now, $n = mk$).  
In the application under consideration in this article, the observations {\it within} a batch are correlated due to the optimization in each batch; see \S \ref{ssec:mrp} for details. 
The sample mean of each batch is computed as
$
	\yb_j = \frac{1}{m} \sum_{i=1}^{m} y^{m(j-1)+i}, j = 1,2, \dots, k,
$
and the overall sample mean $\ybb = \frac{1}{k} \sum_{j=1}^k \yb_j = \frac{1}{n} \sum_{i=1}^n y^i$ provides a point estimator of $\mu$.  
The variance estimator of the batch means, which is $m/n$ times the sample variance, is calculated as
\begin{equation} \label{eq:var}
	\varhat(\ybb) = \frac{m}{n}\frac{1}{k-1} \sum_{j=1}^k \left( \yb_j - \ybb \right)^2.
\end{equation}
The $(1-\alpha)$-level approximate CI is then formed by $\ybb \pm t_{k-1,\alpha/2} \sqrt{\varhat(\ybb)}$, where $t_{k-1,\alpha/2}$ denotes the $1-\alpha/2$ quantile of the Student's t distribution with $k-1$ degrees of freedom.  
Under appropriate conditions, $n\varhat(\ybb)$ is a consistent estimator of $\sigma^2$ \citep{alexo_enc_06}.

@@

Overlapping batch means modifies nonoverlapping batch means by reusing a subset of observations from one batch to form the next batch. 
In this paper, we use the {\it batch nonoverlap parameter} $1 \leq \gamma \leq m$ to denote how much neighboring batches do {\it not} overlap.  
For instance, $\gamma = m$ corresponds to the case of nonoverlapping batches and $\gamma = 1$ corresponds to the maximally overlapping case of \citet{Meketon1984}.  
In general, $\gamma$ can take any integer value between and including $[1,m]$. 
Let $\lfloor \cdot \rfloor$ be the greatest integer that is less than or equal to its argument. 
Here, $b=\nb$ batches are formed given $n$, $m$, and $\gamma$. 
The sample mean of each overlapping batch is calculated similarly, 
$
\yb_j = \frac{1}{m} \sum_{i=1}^m y^{\gamma(j-1) + i},\ \  j = 1, 2, \dots, b, 
$
now taking into account $\gamma$.   
The variance estimator is updated to
\begin{equation} \label{eq:variance_gamma}
	\vartg{\ybb} = \frac{1}{b \left( \tfrac{n}{m} - 1 \right) }  \sum_{j=1}^{b} (\yb_j - \ybb)^2.
\end{equation}

The overlapping variance estimator using $1 \leq \gamma < m$ has (i) nearly the same bias as the standard nonoverlapping variance estimator, and it has (ii) a lower asymptotic variance than its nonoverlapping counterpart. 
Therefore, by simply reusing the data in this fashion one obtains a better variance estimator. 
Observe that with $\gamma=m$ and $n=mk$, (\ref{eq:variance_gamma}) reduces to (\ref{eq:var}).
With $\gamma=1$ (the maximally overlapping case), the denominator in (\ref{eq:variance_gamma}) is slightly different than the one in  \cite{Meketon1984}.
This slightly different denominator makes $\vto{\ybb}$ an unbiased estimator for i.i.d.\ data for all finite $m$ and $n$, with the same asymptotic benefits of nearly the same bias and lower variance relative to its counterpart obtained via nonoverlapping batches \citep{Song1992}.

>>>>

% v2-acmsmall-sample.tex, dated March 6 2012
% This is a sample file for ACM small trim journals
%
% Compilation using 'acmsmall.cls' - version 1.3 (March 2012), Aptara Inc.
% (c) 2010 Association for Computing Machinery (ACM)
%
% Questions/Suggestions/Feedback should be addressed to => "acmtexsupport@aptaracorp.com".
% Users can also go through the FAQs available on the journal's submission webpage.
%
% Steps to compile: latex, bibtex, latex latex
%
% For tracking purposes => this is v1.3 - March 2012

\documentclass[prodmode,acmtomacs]{acmsmall} % Aptara syntax

% Package to generate and customize Algorithm as per ACM style
\usepackage[ruled]{algorithm2e}
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

% Metadata Information   %%% TO BE UPDATED!
\acmVolume{9}
\acmNumber{4}
\acmArticle{39}
\acmYear{2010}
\acmMonth{3}

% User added packages
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{subfig}

% User added commands
% Frequently used general mathematics
\newcommand{\R}{{\mathbb{R}}}
\newcommand{\Rp}{\R^+}
\newcommand{\Z}{{\mathbb{Z}}}
\newcommand{\Zp}{\Z^+}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\N}{\mathbb{N}}

@@

% Document starts
\begin{document}

% Page heads
\markboth{D. Love and G. Bayraksan}{Overlapping Batches for the Assessment of Solution Quality in Stochastic Programs}

% Title portion
\title{Overlapping Batches for the Assessment of Solution Quality in Stochastic Programs}
\author{DAVID LOVE
\affil{University of Arizona}
G\"{U}Z\.{I}N BAYRAKSAN
\affil{The Ohio State University}
}

\begin{abstract}
Overlapping Batch Means (OBM) has long been used in simulation as a method of reusing data to generate variance estimators with asymptotically lower variance.
In this article, we apply the OBM method to stochastic programming by formulating a variant of the multiple replications procedure used for assessing solution quality.
We give conditions under which the resulting optimality gap point estimators are strongly consistent, the optimality gap interval estimators are asymptotically valid, and the OBM variance estimators for optimality gap have asymptotically lower variances relative to their nonoverlapping counterparts \cite{Meketon1984,Welch1987}.
We investigate computational efficiency, a combined measure of variance and computation time, providing guidelines on the degree of overlap.  
Numerical experiments on several test problems are presented, examining the small-sample behavior and the empirical computational efficiency of the overlapping batches method in this context.  
\end{abstract}

\category{G.1.6}{Numerical Analysis}{Optimization}[Stochastic Programming]
\category{I.6.8}{Simulation and Modeling}{Types of Simulation}[Monte Carlo]
\category{I.6.1}{Simulation and Modeling}{Simulation Theory}

\terms{Theory, Algorithms, Measurement}

\keywords{Sample average approximation (SAA); Overlapping batch means;   Stochastic programming; Optimality gap estimation}

\acmformat{David Love and G\"{u}zin Bayraksan, 2014. Overlapping batches for the assessment of solution quality in stochastic programs.}
% At a minimum you need to supply the author names, year and a title.
% IMPORTANT:
% Full first names whenever they are known, surname last, followed by a period.
% In the case of two authors, 'and' is placed between them.
% In the case of three or more authors, the serial comma is used, that is, all author names
% except the last one but including the penultimate author's name are followed by a comma,
% and then 'and' is placed before the final author's name.
% If only first and middle initials are known, then each initial
% is followed by a period and they are separated by a space.
% The remaining information (journal title, volume, article number, date, etc.) is 'auto-generated'.

\begin{bottomstuff}
This work is supported in part by the National Science Foundation, under
grant CMMI-1345626 and grant DMS-0602173.

Author's addresses: D. Love, (Current address) American Express, New York, NY;
G. Bayraksan, Department of Integrated Systems Engineering, the Ohio State University, Columbus, OH; email: bayraksan.1@osu.edu.
\end{bottomstuff}

@@

@@

The Overlapping Batch Means (OBM) method, proposed by \citeN{Meketon1984}, is commonly used in simulation to obtain variance estimators having variances that are comparatively lower than their nonoverlapped counterparts. 
In the context of steady-state simulation, OBM is often used to estimate the variance of the sample mean, which is itself an estimator for the mean. 
\citeN{SAH90} use overlapping estimators to estimate the variance of estimators for non-means. 
More recently, overlapping estimators based on standardized time series (as opposed to batch means) have been evaluated [\citeNP{Alexopoulos01012007};\citeyearNP{Alexopoulos2007}]; see also \citeN{Meterelliyoz_etal_12}.
To the best of our knowledge, OBM has never before been studied in a stochastic optimization context. 
In this article, we show that the overlapping batches achieve nearly the same bias and lower variances relative to their nonoverlapped counterparts when used in a stochastic optimization setting.

A natural setting where batch means are used in stochastic optimization (especially in stochastic programming) is for assessing solution quality. 
This typically involves generating point and interval estimators of the optimality gap of a given candidate solution using Monte Carlo simulation. 
Because optimality gap estimators can be difficult to analyze statistically, a common approach is to use nonoverlapping batches of data to generate a number of optimality gap estimators. 
These individual estimators are then used to form point and interval estimators applying the nonoverlapping batch means method.
This is referred to as the Multiple Replications Procedure (MRP) and was initially discussed in \citeN{Mak1999}.   
In this article, we formulate an overlapping batches version of MRP.

@@

We wish to assess the quality of a candidate solution $\xh \in X$, which may have been found in any way.
Assessing solution quality is important in practice because as (\ref{eq:sto_prog}) typically cannot be solved exactly, one only has an approximate solution $\xh$ without verification of its quality.
Assessing solution quality is also a critical component of stopping criteria in algorithms designed to (approximately) solve (\ref{eq:sto_prog}).
We define the quality of $\xh$ by its optimality gap, $\e{f(\xh,\xit)} - \zs$. 
The lower the optimality gap, the higher the quality of the solution; and a zero optimality gap implies that the solution is optimal. 
We assume that the candidate solution $\xh$ is given as input to our procedures, and hence fixed, throughout the article.

The optimality gap $\e{f(\xh,\xit)} - \zs$ often cannot be calculated exactly. First, for a given candidate solution $\xh \in X$, evaluation of $\e{f(\xh,\xit)}$ typically involves a difficult multidimensional integral. 
Second, we do not know $\zs$. 
In many fields of optimization, the second problem is alleviated by obtaining lower bounds on $\zs$ through relaxations such as integrality, Lagrangian, or semidefinite relaxations.
As a result, it is common to evaluate an upper bound on the optimality gap. 
This motivates estimating the upper bound on the optimality gap in stochastic programming by using Monte Carlo simulation.
Given a sample size $m$, an upper bound on the optimality gap can be obtained by \[
\e{f(\xh,\xit)} - \zs \leq \e{f(\xh,\xit)} - \e{\zs_m}
\]
due to the inequality $\e{\zs_m} \leq \zs$ \cite{Mak1999,norkin_pflug_ruszczynski_98}.
This bound improves as the sample size increases; that is, $\e{\zs_m} \leq \e{\zs_{m+1}} \leq \zs$.  
A straightforward estimate of $\e{f(\xh,\xit)}$ is the sample mean, $\frac{1}{m} \sum_{i=1}^m f(\xh,\xiti)$. 
Instead of $\e{\zs_m}$, we can simply use $\zs_m$.  
The resulting point estimator of the optimality gap is then 
\[
\frac{1}{m} \sum_{i=1}^m f(\xh,\xiti) - \zs_m.
\]
Here, we assume the same observations $\xit^1, \dots, \xit^m$ are used in both terms.

Computing the above optimality gap estimate involves solving the optimization problem (\ref{eq:sto_prog_m}) to obtain a lower bound estimator of $\zs$, $\zs_m$, which complicates the statistical analysis. 
As mentioned above, to enable statistical inference, the MRP of \citeN{Mak1999} generates $k$ independent estimators, each using sample size $m$ ($k$ nonoverlapping batches of size $m$), and averages them to obtain a point estimator.  
The sample variance of these estimators is used to form confidence intervals (CIs) for the optimality gap (we further review MRP in \S \ref{ssec:mrp}).  
An advantage of MRP is its applicability to a wide range of problems.  
With i.i.d.\ sampling, $f$ can be linear or nonlinear, $\X$ can include integrality constraints or not.  
It is also easy to implement; thus, MRP has been applied to a wide variety of problems, including finance, supply chain network design, etc.; see, e.g., \citeN{bertocchi_etal_99,janjarassuk_linderoth_08,santoso_ahmed_etal_05}.  
Recently, the approach of (nonoverlapping) batching has been used for assessing solution quality of stochastic programs with finitely many expected value \cite{wang_ahmed_08} and stochastic dominance \cite{hu2010sample} constraints.

@@

The contributions of this article are as follows: 
\begin{enumerate}
\item We formulate an overlapping batch means variant of MRP, where the degree of overlap can be chosen by the user.  
This application of OBM is novel because each batch mean estimator includes an internal minimization, which has not been studied in the literature.

\item We provide conditions under which the resulting optimality gap estimators exhibit reduced variance when compared to their counterparts obtained via nonoverlapping batches.
We also show that the resulting point estimators are strongly consistent and that the interval estimators are asymptotically valid. 
These results require new proofs because of the presence of the optimization operator.

\item We examine small-sample behavior of the estimators through numerical examples. 
Our experiments indicate that the use of overlapping estimators results in gap estimators achieving lower variance than their nonoverlapping counterparts, along with comparable bias and interval coverage---for both small- and large-sample cases.

\item As the degree of overlap increases, the number of batches---and optimization problems solved---increases because we need to solve a sampling problem per batch. 
At the same time, variances are lowered, and when warm starting is available, solution time per batch can decrease, yielding efficiency gains. 
We explore this trade-off between computation time and lower variances.
Our results indicate that when warm starting is effective, overlapping estimators are computationally more efficient than those formed via nonoverlapping batches.
Based on this study, we provide guidelines on selecting the degree of overlap and demonstrate efficiency on a newsvendor problem. 
\end{enumerate}

An earlier version of the article appeared in \citeN{love2011overlapping}. 
Most of the theoretical results in the current version of the article are new, including almost sure convergence and asymptotically lower variances of the overlapping variance estimators, and asymptotic validity of the interval estimators. 
These results are achieved through a stronger assumption on the rate of convergence of the sampling problem's solution to the unique optimal solution, which is satisfied by several classes of (\ref{eq:sto_prog}); see \S \ref{subsec:assumptions}.
We also present additional computational results, including those from a large-scale problem, and study computational efficiency of the overlapping MRP method.

The rest of the article is organized as follows.  
In the next section, we review relevant background information.  
We first discuss nonoverlapping and overlapping batch means in \S \ref{ssec:obm} and then briefly go over MRP in \S \ref{ssec:mrp}.  
In \S \ref{sec:omrp}, we present the Overlapping Multiple Replications Procedure (OMRP) with a parametrized degree of overlap.  
In \S \ref{sec:theory}, we show strong consistency of the OMRP point estimators, asymptotic validity of the OMRP CIs, and establish that the OMRP variance estimators have lower variances than their nonoverlapping MRP counterparts.
In \S \ref{sec:comp}, we test the performance of OMRP on four problems and study computational efficiency. 
We end in \S \ref{sec:concl} with a summary and conclusions.

@@

Consider a covariance stationary stochastic process $y^1, y^2, \dots$ with mean $\mu$ and variance parameter $0<\sigma^2<\infty$.  
The variance parameter is the sum of covariances at all lags, $\sigma^2=\mathrm{Var} \left(y^1 \right)+2\sum_{j=1}^{\infty}\mathrm{Cov} \left(y^1,y^{1+j}\right)$. 
If the variance parameter exists, then  $\sigma^2=\lim_{n\rightarrow \infty}n\mathrm{Var} \left(\ybb\right)$, where $\ybb=\frac{1}{n}\sum_{i=1}^{n} y^{i}$.
The task is to estimate $\mu$ given some realization of the univariate stochastic process $\y = (y^1, y^2, \dots, y^n)$.  
Typically, this involves forming a CI of the form $[L_\alpha(\y), U_\alpha(\y)]$ for a given level of significance $\alpha$ such that $\p{L_\alpha(\y) \leq \mu \leq U_\alpha(\y)} = 1 - \alpha$ or that this probability is approximately $1 - \alpha$ for large enough sample sizes.  
The usual estimator for $\mu$ is the sample mean $\ybb$, which is an unbiased estimator even in the presence of correlated data.  
However, the usual (sample) variance estimator $\frac{1}{n-1} \sum_{i=1}^n (y^i - \ybb)^2$ can be inappropriate to use in the presence of correlated data, resulting in inaccurate interval estimators.  
To overcome this difficulty, various variance estimators have been proposed in the simulation literature; see, e.g., \citeN{law_15}.  
We briefly review two of these variance estimators: nonoverlapping and overlapping batch means.

@@

Nonoverlapping batch means takes $n$ observations $y^1, \dots, y^n$ of the stochastic process and splits them into $k$ batches of size $m$, where $k = \frac{n}{m}$ (for simplicity, assume, for now, $n = mk$).  
In the application under consideration in this article, the observations {\it within} a batch are correlated due to the optimization in each batch; see \S \ref{ssec:mrp} for details. 
The sample mean of each batch is computed as
$
	\yb_j = \frac{1}{m} \sum_{i=1}^{m} y^{m(j-1)+i}, j = 1,2, \dots, k,
$
and the overall sample mean $\ybb = \frac{1}{k} \sum_{j=1}^k \yb_j = \frac{1}{n} \sum_{i=1}^n y^i$ provides a point estimator of $\mu$.  
The variance estimator of the batch means, which is $m/n$ times the sample variance, is calculated as
\begin{equation} 
	\label{eq:var}
	\varhat(\ybb) = \frac{m}{n}\frac{1}{k-1} \sum_{j=1}^k \left( \yb_j - \ybb \right)^2.
\end{equation}
The $(1-\alpha)$-level approximate CI is then formed by $\ybb \pm t_{k-1,\alpha/2} \sqrt{\varhat(\ybb)}$, where $t_{k-1,\alpha/2}$ denotes the $1-\alpha/2$ quantile of the Student's t distribution with $k-1$ degrees of freedom.  
Under appropriate conditions, $n\varhat(\ybb)$ is a consistent estimator of $\sigma^2$ \cite{alexo_enc_06}.

@@

Overlapping batch means modifies nonoverlapping batch means by reusing a subset of observations from one batch to form the next batch. 
In this article, we use the {\it batch nonoverlap parameter} $1 \leq \gamma \leq m$ to denote how much neighboring batches do {\it not} overlap.  
For instance, $\gamma = m$ corresponds to the case of nonoverlapping batches and $\gamma = 1$ corresponds to the maximally overlapping case of \citeN{Meketon1984}.  
In general, $\gamma$ can take any integer value between and including $[1,m]$. 
Let $\lfloor \cdot \rfloor$ be the greatest integer that is less than or equal to its argument. 
Here, $b=\nb$ batches are formed given $n$, $m$, and $\gamma$. 
The sample mean of each overlapping batch is calculated similarly, 
$
\yb_j = \frac{1}{m} \sum_{i=1}^m y^{\gamma(j-1) + i},\ \  j = 1, 2, \dots, b, 
$
now taking into account $\gamma$.   
The variance estimator is updated to
\begin{equation} 
	\label{eq:variance_gamma}
	\vartg{\ybb} = \frac{1}{b \left( \tfrac{n}{m} - 1 \right) }  \sum_{j=1}^{b} (\yb_j - \ybb)^2.
\end{equation}

The overlapping variance estimator using $1 \leq \gamma < m$ has (i) nearly the same bias as the standard nonoverlapping variance estimator, and it has (ii) a lower asymptotic variance than its nonoverlapping counterpart. 
Therefore, by simply reusing the data in this fashion one obtains a better variance estimator. 
Observe that with $\gamma=m$ and $n=mk$, (\ref{eq:variance_gamma}) reduces to (\ref{eq:var}).
With $\gamma=1$ (the maximally overlapping case), the denominator in (\ref{eq:variance_gamma}) is slightly different than the one in  \citeN{Meketon1984}.
This slightly different denominator makes $\vto{\ybb}$ an unbiased estimator for i.i.d.\ data for all finite $m$ and $n$, with the same asymptotic benefits of nearly the same bias and lower variance relative to its counterpart obtained via nonoverlapping batches \cite{Song1992}.

The amount of asymptotic reduction in $\var{\vartg{\ybb}}$ for a given $1\leq \gamma< m$ relative to its nonoverlapping counterpart depends on the asymptotic ratio of $\gamma/m$, which we denote by $\gammab$. 
The maximally overlapping case ($\gamma = 1$) corresponds to $\gammab=0$, and the nonoverlapping case corresponds to $\gammab=1$. 
Damerdji~\shortcite{damerdji1994strong,damerdji1995mean} studied batching structures that guarantee almost sure and mean-square consistency of $\vartg{\ybb}$ in the nonoverlapping and maximally overlapping cases. 
Under appropriate conditions, the results of \citeN{damerdji1995mean}, along with those of \citeN{Meketon1984} and \citeN{Welch1987}, establish that 
\begin{equation}
	\label{eq:varvarred}
	\frac{n}{m}\var{n \vartg{\ybb}} \rightarrow \beta_{\gammab} \sigma^4
\end{equation}
as the batch size $m$ and then the number of batches $n/m$ tend to infinity; see also \citeN{Alexopoulos2007}. 
In (\ref{eq:varvarred}), $\beta_1 = 2$ for nonoverlapping batches and $\beta_0 = \frac{4}{3}$ for maximally overlapping batches.  
More generally, $\beta_{\gammab}$ is twice the value of 
\begin{equation} 
	\label{eq:var_reduct_formula}
	\gammab \left( 1 + 2 \sum_{j=1}^{\cogb-1} (1-j\gammab)^2 \right),
\end{equation}
where $\lceil \cdot \rceil$ denotes the smallest integer that is greater than or equal to its argument \cite{Welch1987}.
If $\ogb$ is an integer, (\ref{eq:var_reduct_formula}) simplifies to $\frac{2+\gammab^2}{3}$.
This means that the maximally overlapping variance estimator has a variance that is two-thirds ($66.67\%$) of the nonoverlapping estimator.
As the degree of overlap decreases, the variance increases. 
When only half of the observations overlap 
($\gammab = 1/2$), for instance, the variance is 75\% of the nonoverlapping counterpart.
